---
title: "Getting started with flexdashboard"
author: "Kerry Cella"
date: "22 April 2018"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)

cc_data<- read.csv("Childcare_provider_data_Ofsted_Dec2017.csv",
                   header = TRUE )
```

## Introduction

This session will cover how to get started building flexdashboard. 

<br>

A flexdashboard is created using R Markdown and allows you to create interactive and visually appealing dashboard easily.

<br>

A complementary getting started guide is [here](https://rmarkdown.rstudio.com/flexdashboard/)


## Why use a flexdashboard

- They are relatively simple to build

- They provide a level of interactivity with html widgets

- They look good

- You don't need Shiny! (but you can use it)

## Get started

First install the flexdashboard package
```{r cars, eval= FALSE, echo =TRUE}
install.packages("flexdashboard")
```

Then in R Studio do the following:

<br>

**File -> New File -> R Markdown... -> From Template -> Flex Dashboard**

## Layout

<font size  = 4>

A layout can be divided into columns or rows. You can decide the orientation in the YAML at the top of your markdown document.

```{r, eval = FALSE, echo=TRUE}
title: "Row Orientation"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
```

To create a column in the main area under the YAML:

```{r eval =FALSE, echo=TRUE}
Column
-------------------------------------------------
```

To create a row

```{r eval =FALSE, echo=TRUE}
Row
-------------------------------------------------
```

</font>

## Theme

You can also choose the theme within the YAML.

```{r, eval = FALSE, echo=TRUE}
title: "Row Orientation"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    theme: cerulean
```

A list of avalable themes is [here](https://rmarkdown.rstudio.com/flexdashboard/using.html#appearance)

## Insert a chart

<font size = 4>

To insert a chart you add a header as follows:

```{r, echo= TRUE}
### Chart A
```

You can then insert a chart below that within an R chunk

```{r, warning=FALSE, message= FALSE, fig.height=3, fig.width= 7}
my_graph<- ggplot(data = cc_data) +
  geom_bar(aes(x = Provider.Type, fill = Provider.Type), stat = "count")+
  labs(title = "Chidcare market Ofsted 2017")+
  theme_bw()+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

my_graph

```

</font>

## Add another page

Adding another page is simple. You delineate as follows:

```{r, echo = TRUE, eval= FALSE}

Front Page
=============================================

Row
---------------------------------------------
### Chart 1

Row
---------------------------------------------
  
Page 2
=============================================

```

## Insert a valuebox

<font size  = 4>

You can insert a value box in a similar way to how you insert a chart

    ### My favourite number
    ```{r}`r ''`
    valueBox(10, icon = "fa-ceritifcate")
     ```
You can also customise the appareance and the icon:

      ```{r}`r ''`
      most_places<- max(cc_data$Registered.places)
      valueBox(most_places, icon = "fa-bullhorn", color = "green")
      ```

The icon options are explained [here](https://rmarkdown.rstudio.com/flexdashboard/using.html#value_boxes)

</font>

## Insert a sidebar

```{r, eval= FALSE, echo= TRUE}

Inputs {.sidebar}
-------------------------------------

Row  
-------------------------------------  
```

## General lesson learned

- Ask for feedback from your users

- You might find it's the simplest features that offer most value

- R markdown / flex - great way to start learning html / css
